# Load libraries

```{r}
rm(list=ls())

library(AER)
library(car)
library(FactoMineR)
```


#Load data and take a look

```{r}
file_path = "C:/Users/jordi.catafal/Desktop/heart.csv"
df = read.csv(file_path, header = T)
head(df)
summary(df)
```

#Categorical data: change numeric value to categorical

```{r}
df$sex[which(df$sex==0)] <- "Female"
df$sex[which(df$sex==1)] <- "Male"
df$sex <- as.factor(df$sex)

df$cp[which(df$cp==0)] <- "Typical Angina"
df$cp[which(df$cp==1)] <- "Atypical Angina"
df$cp[which(df$cp==2)] <- "Non-anginal Pain"
df$cp[which(df$cp==3)] <- "Asymptomatic"
df$cp <- as.factor(df$cp)

df$fbs[which(df$fbs==0)] <- "<= 120 mg/dL"
df$fbs[which(df$fbs==1)] <- "> 120 mg/dL"
df$fbs <- as.factor(df$fbs)

df$restecg[which(df$restecg==0)] <- "Normal"
df$restecg[which(df$restecg==1)] <- "Abnormality"
df$restecg[which(df$restecg==2)] <- "Hypertrophy"
df$restecg <- as.factor(df$restecg)

df$exang[which(df$exang==0)] <- "No"
df$exang[which(df$exang==1)] <- "Yes"
df$exang <- as.factor(df$exang)

df$slope[which(df$slope==0)] <- "Upsloping"
df$slope[which(df$slope==1)] <- "Flat"
df$slope[which(df$slope==2)] <- "Downsloping"
df$slope <- as.factor(df$slope)

df$thal[which(df$thal==0)] <- "Normal"
df$thal[which(df$thal==1)] <- "Fixed Defect"
df$thal[which(df$thal==2)] <- "Reversible Defect"
df$thal <- as.factor(df$thal)
```


```{r}
hist(df$age)

plot(df$sex)

sapply(df, class)

df$age <- as.numeric(df$age)
df$trestbps <- as.numeric(df$trestbps)
df$chol <- as.numeric(df$chol)
df$thalach <- as.numeric(df$thalach)
df$oldpeak <- as.numeric(df$oldpeak)

sapply(df, class)
```


# Detection of errors

```{r}
summary(df)

df$thal[which(df$thal==3)] <- NA

miss_val = sum(is.na(df))

miss_val

which(is.na(df))



plot(df$thalach)
hist(df$oldpeak)

summary(df)
sapply(df, class)
```

##kable -> taula

# Univariate detection

```{r}
outliers <- function(column, name){
  
  sumlist <- summary(column)
  q1 <- sumlist[2] 
  q3 <- sumlist[5]
  
  
  boxplot(column, main = paste("Boxplot ", name), col = "orange", horizontal = T)
  
  # IQR calculation
  iqr <- q3 - q1 
  
  # Mild inferior limit:
  mild_inf_lim <- sumlist[2]-1.5*iqr
  # Extreme inferior limit:
  extreme_inf_lim <- sumlist[2]-3*iqr
  abline(v=mild_inf_lim, col = "red", lty = 2)
  abline(v=extreme_inf_lim, col = "red",  lty = 2, lwd = 2)
  mild_sup_lim <- sumlist[5]+1.5*iqr
  extreme_sup_lim <- sumlist[5]+3*iqr
  abline(v=mild_sup_lim, col = "red", lty = 2)
  abline(v=extreme_sup_lim, col = "red",  lty = 2, lwd = 2)
  
}
```

```{r}
outliers(df$age, "age")
outliers(df$trestbps, "trestbps")
outliers(df$chol, "chol")
outliers(df$thalach, "thalach")
outliers(df$oldpeak, "oldpeak")
outliers(df$ca, "ca")
outliers(df$target, "target")
```

```{r}
FindMildOutliers <- function(data) {
  lowerq = quantile(data)[2]
  upperq = quantile(data)[4]
  iqr = upperq - lowerq #Or use IQR(data)
  # we identify extreme outliers
  extreme.threshold.upper = (iqr * 1.5) + upperq
  extreme.threshold.lower = lowerq - (iqr * 1.5)
  result <- which(data > extreme.threshold.upper | data < extreme.threshold.lower)
}

FindExtremeOutliers <- function(data) {
  lowerq = quantile(data)[2]
  upperq = quantile(data)[4]
  iqr = upperq - lowerq #Or use IQR(data)
  # we identify extreme outliers
  extreme.threshold.upper = (iqr * 3) + upperq
  extreme.threshold.lower = lowerq - (iqr * 3)
  result <- which(data > extreme.threshold.upper | data < extreme.threshold.lower)
}


FindMissingValues <- function(data) {
  result <- which(sum(is.na(data)) > 0)
}
```



## Finding all Extreme Outliers by Columns
```{r}
df2 <- Filter(is.numeric, df)
pos <- lapply(df2, FindExtremeOutliers)
lengths(pos)
num_outliers = length(pos)
```


## Finding all Mild Outliers by Columns
```{r}
df2 <- Filter(is.numeric, df)
pos <- lapply(df2, FindMildOutliers)
lengths(pos)
```

## Finding all Missing Values by Individuals
```{r}
pos <- apply(df, 1, FindMissingValues)
pos <- which(pos > 0)
pos
```

## Finding all Extreme Outliers by Individuals
```{r}
df2 <- Filter(is.numeric, df)
pos <- apply(df2, 1, FindExtremeOutliers)
pos <- which(pos > 0)
pos
```

## Finding all Mild Outliers by Individuals
```{r}
df2 <- Filter(is.numeric, df)
pos <- apply(df2, 1, FindMildOutliers)
pos <- which(pos > 0)
pos
length(pos)
```

```{r}
boxplot(df$chol)
boxplot(df$fbs)
boxplot(df$restecg)
```


## Create variable adding the total number missing values, outliers and errors.

```{r}
num_outliers_miss_errors = miss_val + num_outliers
num_outliers_miss_errors
```


## Compute the correlation with all other variables. Rank these variables according the correlation
```{r}
library(FactoMineR)
```



### Profiling

```{r}
# Categorical output:
library(FactoMineR)
summary(df$target)
res.condes <- condes(df, 14, proba = 0.50)
res.condes$quanti  # Global association to numeric variables
res.condes$quali # Partial association of numeric variables to levels of outcome factor
res.condes$category  # Partial association to significative levels in factors
```



